---
title: "globaltrends SPRI"
subtitle: " A Google Trends-Based Socio-Political Risk Measure"
format: gfm
---

We propose a concept of grassroots socio-political risk (SPR) and provide the
**globaltrends SPRI** as a measure. The **globaltrends SPRI** is based on
individual-level issue salience using data from
[Google Trends](https://trends.google.com/trends/). Our concept provides a
bottom-up complement to established measures that focus more on political
institutional players and the institutional constraints they face. The
**globaltrends SPRI** can serve to predict institutional change in a country
initiated by a popular majority of individuals. In contrast to existing
measures, the **globaltrends SPRI** measure captures issue salience directly
among individuals rather than media or analyst reports. As such, it is uniquely
representative of socio-political sentiment, timely, and widely available. Our
concept and measure offer novel opportunities for international business
researchers. Refer to the accompanying paper (available from the authors) for
further information on the index construction.

The current version of the **globaltrends SPRI** dataset can be accessed in
differnet file formats:

+ [PARQUET](https://github.com/ha-pu/globaltrends_spri/blob/main/data_spri.parquet)
+ [RDS](https://github.com/ha-pu/globaltrends_spri/blob/main/data_spri.rds)
+ [XLSX](https://github.com/ha-pu/globaltrends_spri/blob/main/data_spri.xlsx)

Cite any usage of the **globaltrends SPRI** as:  
Puhr, H., & Müllner, J. (2023). Vox Populi, Vox Dei: A Concept and Measure for
Grassroots Socio-Political Risk Using Google Trends.
*Journal of International Management*, 30(2): 101096.
[https://doi.org/10.1016/j.intman.2023.101096](https://doi.org/10.1016/j.intman.2023.101096)

The **globaltrends SPRI** is part of the
[globaltrends project](https://github.com/ha-pu/globaltrends).

* Last update of the data: 2026-01-07
* Coverage of the data: 2010-2025

## Development of Global SPRI

```{r, packages}
#| echo: false
#| mesage: false
#| warning: false

library(glue)
library(lubridate)
library(maps)
library(readxl)
library(tidyverse)
library(writexl)
```

```{r, load_data}
#| echo: false
#| mesage: false
#| warning: false

category_mean <- read_rds("data/category_base.rds")
category_global <- read_rds("data/category_global.rds")
category_internet <- read_rds("data/category_internet.rds")

data_wdi <- read_rds("data/data_wdi.rds")

lst_countries <- read_xlsx("input/spri_countries.xlsx")

year <- read_lines("input/new_year.txt")
```

```{r, combine_data}
#| echo: false
#| mesage: false
#| warning: false

data_wdi <- lst_countries %>%
  inner_join(data_wdi, by = "iso2c", multiple = "all") %>%
  select(country = country.x, year, gdp_share)

data <- category_mean %>%
  group_by(location, date, year) %>%
  summarise(spri = mean(spri), .groups = "drop") %>%
  mutate(category = "Total") %>%
  bind_rows(category_mean) %>%
  mutate(spri = spri * 100)

data <- inner_join(data, lst_countries, by = c("location" = "iso2c"))
  
data_global <- category_global %>%
  group_by(date, year) %>%
  summarise(spri = mean(spri), .groups = "drop") %>%
  mutate(category = "Total") %>%
  bind_rows(category_global) %>%
  mutate(spri = spri * 100)

data_internet <- category_internet %>%
  group_by(date, year) %>%
  summarise(spri = mean(spri), .groups = "drop") %>%
  mutate(category = "Total") %>%
  bind_rows(category_internet) %>%
  mutate(spri = spri * 100)

data_global <- left_join(data_global, data_internet, by = c("date", "year", "category"))
```

```{r, plot_global_spri0}
#| echo: false
#| mesage: false
#| warning: false

categories <- c("Total", "Economy", "Government", "Security", "Society")
year_min <- year(min(data_global$date))
year_max <- year(max(data_global$date))
year_seq <- seq(year_min, year_max)
date_seq <- dmy(paste(31, 12, year_seq, sep = "-"))
```

```{r, plot_global_spri1}
#| echo: false
#| mesage: false
#| warning: false

.x <- 1
data_global %>%
    filter(category == categories[[.x]]) %>%
    ggplot() +
    geom_line(aes(x = date, y = spri.x), colour = "darkred") +
    geom_line(aes(x = date, y = spri.y), colour = "darkblue") +
    scale_x_date(minor_breaks = date_seq) +
    labs(
      x = NULL,
      y = glue("Grassroots SPRI - {categories[[.x]]}"),
      title = glue("Grassroots SPRI - {categories[[.x]]}")
    ) +
    theme_bw()
```

```{r, plot_global_spri2}
#| echo: false
#| mesage: false
#| warning: false

.x <- 2
data_global %>%
    filter(category == categories[[.x]]) %>%
    ggplot() +
    geom_line(aes(x = date, y = spri.x), colour = "darkred") +
    geom_line(aes(x = date, y = spri.y), colour = "darkblue") +
    scale_x_date(minor_breaks = date_seq) +
    labs(
      x = NULL,
      y = glue("Grassroots SPRI - {categories[[.x]]}"),
      title = glue("Grassroots SPRI - {categories[[.x]]}")
    ) +
    theme_bw()
```

```{r, plot_global_spri3}
#| echo: false
#| mesage: false
#| warning: false

.x <- 3
data_global %>%
    filter(category == categories[[.x]]) %>%
    ggplot() +
    geom_line(aes(x = date, y = spri.x), colour = "darkred") +
    geom_line(aes(x = date, y = spri.y), colour = "darkblue") +
    scale_x_date(minor_breaks = date_seq) +
    labs(
      x = NULL,
      y = glue("Grassroots SPRI - {categories[[.x]]}"),
      title = glue("Grassroots SPRI - {categories[[.x]]}")
    ) +
    theme_bw()
```

```{r, plot_global_spri4}
#| echo: false
#| mesage: false
#| warning: false

.x <- 4
data_global %>%
    filter(category == categories[[.x]]) %>%
    ggplot() +
    geom_line(aes(x = date, y = spri.x), colour = "darkred") +
    geom_line(aes(x = date, y = spri.y), colour = "darkblue") +
    scale_x_date(minor_breaks = date_seq) +
    labs(
      x = NULL,
      y = glue("Grassroots SPRI - {categories[[.x]]}"),
      title = glue("Grassroots SPRI - {categories[[.x]]}")
    ) +
    theme_bw()
```

```{r, plot_global_spri5}
#| echo: false
#| mesage: false
#| warning: false

.x <- 5
data_global %>%
    filter(category == categories[[.x]]) %>%
    ggplot() +
    geom_line(aes(x = date, y = spri.x), colour = "darkred") +
    geom_line(aes(x = date, y = spri.y), colour = "darkblue") +
    scale_x_date(minor_breaks = date_seq) +
    labs(
      x = NULL,
      y = glue("Grassroots SPRI - {categories[[.x]]}"),
      title = glue("Grassroots SPRI - {categories[[.x]]}")
    ) +
    theme_bw()
```

## Global SPRI 2010 vs. 2025

```{r, plot_map}
#| echo: false
#| mesage: false
#| warning: false

world_data <- data %>%
  group_by(year, country, category) %>%
  summarise(spri = mean(spri), .groups = "drop")

world_data <- map_data("world") %>%
  filter(region != "Antarctica") %>%
  fortify() %>%
  left_join(world_data, by = c("region" = "country"), relationship = "many-to-many")

mean_global <- data_global %>%
  filter(category == "Total") %>%
  summarise(mean_global = log(mean(spri.x))) %>%
  pull(mean_global)
```

```{r, plot_map_old}
#| echo: false
#| mesage: false
#| warning: false

world_old <- world_data %>%
  filter((is.na(year) | year == 2010) & (is.na(category) | category == "Total")) %>%
  mutate(spri = na_if(spri, 0)) %>%
  mutate(spri = log(spri))

ggplot(data = world_old) +
  geom_map(
    map = world_old,
    aes(
      x = long,
      y = lat,
      group = group,
      map_id = region,
      fill = spri
    ),
    colour = "black",
    linewidth = 0.5
  ) +
  scale_y_continuous(breaks = c()) +
  scale_x_continuous(breaks = c()) +
  scale_fill_gradient2(midpoint = mean_global, low = "darkgreen", mid = "darkorange", high = "darkred") +
  labs(
    fill = "Grassroots SPRI Total",
    title = "Grassroots SPRI Total: 2010",
    x = NULL,
    y = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r, plot_map_new}
#| echo: false
#| mesage: false
#| warning: false

world_new <- world_data %>%
  filter((is.na(year) | year == year) & (is.na(category) | category == "Total")) %>%
  mutate(spri = na_if(spri, 0)) %>%
  mutate(spri = log(spri))

ggplot(data = world_new) +
  geom_map(
    map = world_new,
    aes(
      x = long,
      y = lat,
      group = group,
      map_id = region,
      fill = spri
    ),
    colour = "black",
    linewidth = 0.5
  ) +
  scale_y_continuous(breaks = c()) +
  scale_x_continuous(breaks = c()) +
  scale_fill_gradient2(midpoint = mean_global, low = "darkgreen", mid = "darkorange", high = "darkred") +
  labs(
    fill = "Grassroots SPRI Total",
    title = glue("Grassroots SPRI Total: {year}"),
    x = NULL,
    y = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```